### 2022-07-21 draft.tex
Deploying to gh-pages from @ FrancisBanville/ms_maxent_networks@a05ab93f84c626d14312205bd08c03a04957647d 🚀

We predicted the joint degree distribution of maximum entropy for each
food web in our complete dataset, i.e.~using their numbers of species
and numbers of interactions as state variables. We then sampled one
realization of the degree sequence for each network using the
probabilities given by the joint degree distribution. In
fig.~\ref{fig:joint_dd} (left panels), we show the relationship between
\(k_{out}\) and \(k_{in}\) standardized by the number of species in
their networks, for empirical and maximum entropy joint degree
distributions. We see that our model predicts higher values of
generality and vulnerability compared to empirical food webs
(i.e.~relative values of \(k_{out}\) and \(k_{in}\) closer to \(1\)).
However, plotting the difference between predicted and empirical values
for each species gives a different perspective. The right panel of
fig.~\ref{fig:joint_dd} presents these differences when species are
ordered by their total degree in their network (i.e.~by the sum of their
in and out-degrees). Indeed, our predicted joint degree sequences have
the same number of species as their empirical counterparts, but they are
species agnostic; in other words, instead of predicting a value for each
species directly, we predicted the entire joint degree sequence without
taking into account species' identity. When we associate predictions and
empirical data according to their rank in total degrees, we see that
species predicted to be have a higher generality (number of preys)
generally have a lower vulnerability (number of preys) than what is
observed (and conversely). In fig.~S1, we show how these differences
change when species are instead ordered by their out-degrees (left
panel) and in-degrees (right panel), respectively.

@@

We plotted the Kullback--Leibler divergence between in and out-degree
distributions to compare the symmetry of empirical and maximum entropy
joint degree sequences fig.~\ref{fig:kl_diverg}. As we expected, our
model predicted more similar in-degree and out-degree distributions than
empirical data. However, this difference decreased with connectance
right panel of \ref{fig:kl_diverg}. Overall, this suggests that other
ecological constraints might be needed to account for the asymmetry of
the joint degree distribution, especially for networks with a lower
connectance.

@@

Overall, we found that models based on the joint degree sequence (type
II null and MaxEnt network models) reproduced the structure of empirical
networks much better than the ones based on connectance and the neutral
model of relative abundances. On average, the type II MaxEnt network
model was better at predicting nestedness (\(0.62 \pm 0.08\)) than its
corresponding null model (\(0.73 \pm 0.05\); empirical networks:
\(0.63 \pm 0.09\)), as well as the proportion of cannibal species.
However, the type II null model was better at predicting network
diameter and average maximum similarity between species pairs.
Predictions were similar between both type II models for the maximum
trophic level and the proportion of omnivorous species.

@@

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

@@

@@

>>>>

We predicted the joint degree distribution of maximum entropy for each
food web in our complete dataset, i.e.~using their numbers of species
and numbers of interactions as state variables. We then sampled one
realization of the degree sequence for each network using the
probabilities given by the joint degree distribution. In
fig.~\ref{fig:joint_dd} (left panels), we show the relationship between
\(k_{out}\) and \(k_{in}\) standardized by the number of species in
their networks, for empirical and maximum entropy joint degree
distributions. We see that our model predicts higher values of
generality and vulnerability compared to empirical food webs
(i.e.~relative values of \(k_{out}\) and \(k_{in}\) closer to \(1\)).
However, plotting the difference between predicted and empirical values
for each species gives a different perspective. The right panel of
fig.~\ref{fig:joint_dd} presents these differences when species are
ordered by their total degree in their network (i.e.~by the sum of their
in and out-degrees). Indeed, our predicted joint degree sequences have
the same number of species as their empirical counterparts, but they are
species agnostic; in other words, instead of predicting a value for each
species directly, we predicted the entire joint degree sequence without
taking into account species' identity. The challenge is to associate
predictions with empirical data. If we associate each node according to
their rank in total degrees, we see that species predicted to be have a
higher generality (number of preys) generally have a lower vulnerability
(number of preys) than what is observed (and conversely). This is
because the difference in total degree (\(k_{out} + k_{in}\)) is
minimized when species are ordered by their total degrees in both
predicted and empirical sets (i.e.~the average sum of relative
\(k_{out}\) and \(k_{in}\) is close to \(0\) across all species). This
result shows that the differences between predicted and empirical
degrees (in and out) are low for most species when ordered by their
total degrees and that there is no apparent biases towards in or out
degrees. In fig.~S1, we show how these differences change when species
are instead ordered by their out-degrees (left panel) and in-degrees
(right panel), respectively.

@@

We can describe the shape of the joint degree distribution by comparing
its marginal distributions with one another. Doing so allows us to
evaluate how much the shape of the joint degree distribution of maximum
entropy differs from empirical data. We plotted the Kullback--Leibler
(KL) divergence between in and out-degree sequences to compare the
symmetry of empirical and maximum entropy joint degree sequences
fig.~\ref{fig:kl_diverg}. As we expected, our model predicted more
similar in-degree and out-degree distributions than empirical data
(shown by a lower KL divergence). However, this difference decreased
with connectance right panel of \ref{fig:kl_diverg}. This suggests that
other ecological constraints might be needed to account for the
asymmetry of the joint degree distribution, especially for networks with
a lower connectance. This might be due to the fact that food webs with
low connectance are typically harder to predict than food webs with high
connectance. Indeed, in low connectance systems, what makes two species
interact might be more important for predictions than in high
connectance systems, in which what prevents species from interacting
might be more meaningful. This implies that more ecological information
might be needed in food webs with fewer numbers of interactions because
more nuances determine interactions compared to non-interactions.

@@

Overall, we found that models based on the joint degree sequence (type
II null and MaxEnt network models) reproduced the structure of empirical
networks much better than the ones based on connectance and the neutral
model of relative abundances. On average, the type II MaxEnt network
model was better at predicting nestedness (\(0.62 \pm 0.08\)) than its
corresponding null model (\(0.73 \pm 0.05\); empirical networks:
\(0.63 \pm 0.09\)), as well as the proportion of cannibal species. This
might in part be due because nestedness was calculated using the
spectral radius of the adjacency matrix, which leverages information on
the network itself just like the MaxEnt network models. However, the
type II null model was better at predicting network diameter and average
maximum similarity between species pairs. Predictions were similar
between both type II models for the maximum trophic level and the
proportion of omnivorous species.

@@

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

@@

Our results bring to the forefront the role of the joint degree
distribution in shaping food-web structure. This echoes the work of
Williams (2011), which shows using MaxEnt how the degree distribution
can drive the structure of bipartite networks. Fortuna et al. (2010)
also found that the degree distribution of ecological networks is
associated with some of their emerging properties such as nestedness and
modularity. Network ecologists tend to focus on several emerging
properties of food webs (Delmas et al. 2019), although we show that
food-web structure is very well predicted by the joint degree
distribution. In the light of our results, there is a lot more
ecological information in the difference between properties of empirical
networks and MaxEnt predictions than in these properties themselves.

@@

\hypertarget{conclusion}{%
\subsection{Conclusion}\label{conclusion}}
